---
title: "Hw6"
author: "Trevor Freeland"
date: "May 1, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = F, warning = F)
```

```{r}
library(tidyverse)
library(pander)
```

```{r}
prison <- read.csv("http://math.carleton.edu/Chihara/Stats345/Recidivism.csv")
```

##54

###(A)

$f(y;\lambda) = exp[rlog(\lambda) + (r-1)log(y) - log(\Gamma(r) - y\lambda]$

$f(y;\lambda) = exp[- y\lambda + rlog(\lambda) + (r-1)log(y) - log(\Gamma(r) ]$

If we let $\Theta$ = $\lambda$ then we can see this is from the exponential family.

###(B)

E(a(Y)) = $\frac{-c'(\lambda)}{d'(\lambda)}$

E(a(Y)) = $\frac{\delta}{\delta\lambda} *  \frac{-(rlog(\lambda))}{(-\lambda)}$

E(a(Y)) = $\frac{r}{\lambda}$

Var(a(Y)) = $\frac{b''(\lambda)c'(\lambda) - c''(\lambda)d'(\lambda)}{b'(\lambda)^3}$

Var(a(Y)) = $\frac{0 +\frac{r}{\lambda^2} * -1}{-1}$

Var(a(Y)) = $\frac{r}{\lambda^2}$

##55

###(A)

$f(y;\alpha) = exp[log(\alpha) - (\alpha+1)log(y)]$
If we let $\Theta$ = $\alpha$ then we can see this is from the exponential family.

###(B)

E(log(Y)) = $\frac{-c'(\lambda)}{d'(\lambda)}$

E(log(Y)) = $\frac{-1/\alpha}{-1}$

E(log(Y)) = $\frac{1}{\alpha}$

##56

###(A)

In our table below we look at some of the summary statistics of our data. When looking at this table we can see that under timeserved the maximum time served is 219 months and the mean and median are both under 20, this means we could have a potential outlier or two that we might want to be weary of. Then graphically I looked at a few histograms of the data grouped by whether or not the released prisoner was back in prison. In our age plot we can see the histogram shifts slightly left which indicates that younger aged prisoners may be more likely to return to prisoner which makes some intuitive sense. Our other histogram looks at timeserved and there appears to possibly be a minor shift to the right, indicating that the longer someone serves the more likely they may be to return to prison on average. 

```{r}
prison.small <- prison %>% select(c("backinprison", "alcohol", "drugs", "married", "felon", "priors", "educ", "age.years", "timeserved"))
pander(summary(prison.small))
```

```{r}
ggplot(prison.small, aes(x = age.years)) + geom_histogram() + facet_grid(~backinprison) + ggtitle("Age by Recidivism")
```

```{r}
ggplot(prison.small, aes(x = timeserved)) + geom_histogram() + facet_grid(~backinprison) + ggtitle("Time Served by Recidivism")
```

###(B)

$log(\frac{\hat{p}}{1-\hat{p}}) =  .27 + .44(AlcoholYes) -.39(FelonYes) + .25(DrugsYes) + .13(Priors) -.05(Age) + .02(TimeServed)$

```{r, results='hide'}
prison.lm <- glm(backinprison~alcohol+drugs+married+felon+priors+educ+age.years+timeserved, family = binomial, data = prison.small)
summary(prison.lm)
prison.lm2 <- glm(backinprison~alcohol+drugs+married+felon+priors+age.years+timeserved, family = binomial, data = prison.small)
summary(prison.lm2)
anova(prison.lm,prison.lm2, test="Chisq")

#An anova taking both drugs and married out gives me a p-value of .04 so it implies one of them is not 0. Taking each individually out gives me p-values just larger than .05. I decided that Marriage had the larger p-value so I took marriage out and am leaving in drugs. 

prison.lm3 <- glm(backinprison~alcohol+felon+drugs+priors+age.years+timeserved, family = binomial, data = prison.small)
summary(prison.lm3)
anova(prison.lm3,prison.lm2, test="Chisq")
```

##57

###(A)

The odds ratio for a 20-yr old returning to prison compared to a 30 year old returning to prison is 1.65. This indicates that the odds of a 20-yr old returning to prison is 65% larger than the odds a 30-yr old returns to prison holding everything else constant.

```{r}
odds.ratio <- exp(-.05*20)/exp(-.05*30)
```

###(B)

In our plot we show the different effects of Age between prisoners who had no priors, were not felons, spent 24 months in prison, were not convicted on drug charges and either did or did not have alcohol = 1. THe Red line is the distribution for those who had alcohol = 1, and the blue line is for the group that had alcohol = 0. 

```{r}
#Intercept + 24 months of time served + Age + Alcohol (Red)
fun1 <- function(X){
  Y <- .27  + -.05 * X + .44 + 24*.02
  p <- exp(Y)/(1+exp(Y))
  return(p)
}

#Intercept + 24 months of time served + Age (Blue)
fun2 <- function(X){
  Y <- .27  + -.05 * X  + 24*.02
  p <- exp(Y)/(1+exp(Y))
  return(p)
}
plot(0:100, fun1(0:100), type="l", col="red", ylab = "Probability of Returning to Prison" , xlab="Age (yrs)")
lines(0:100, fun2(0:100), col="blue")
```